---
title: "p8105_hw1_ew2717"
author: "Eric Wang"
output: github_document
---

```{r include = FALSE}

library(tidyverse)
library(ggplot2)

```

## Problem 1

#### Creating a data frame that consists of numeric, logical, character, and factor vectors.

```{r creating data frame}

set.seed(10)

data_table <- tibble(
  my_variable = rnorm(10),
  logical_vec = my_variable > 0,
  ten_char_vector = c("c", "h", "a", "r", "a", "c", "t", "e", "r", "s"),
  three_levels = factor(c("one", "two", "three", "three", "two", "one", "one", "one", "three", "three"))
)

data_table
```

#### Attempting to take the means of the variables

```{r finding mean of the variables}

mean(pull(data_table, my_variable))
mean(pull(data_table, logical_vec))
mean(pull(data_table, ten_char_vector))
mean(pull(data_table, three_levels))

```

The mean of my_variable and logical_vec can be calculated because they are numeric and logical vectors. ten_char_vector and three_levels, however, cannot have their means calculated because they are character and factor vectors.

#### Converting logicial, character, and factor variables using as.numeric without showing the outputs

```{r converting to variables to numeric, message = FALSE, warning = FALSE}

converted_num1 = as.numeric(pull(data_table, logical_vec))
converted_num2 = as.numeric(pull(data_table, ten_char_vector))
converted_num3 = as.numeric(pull(data_table, three_levels))

```

#### Finding the means of the converted variables

```{r finding converted mean}

mean(converted_num1)
mean(converted_num2)
mean(converted_num3)

```

The logical and factor vectors can be converted to numeric values, while the character value cannot be converted. We are unable to convert characters or letters to numbers. For the logical vector, we are able to convert it to numeric because 1 can represent TRUE and 0 can represent FALSE. 

After converting logical_vec, ten_char_vector, and three_levels to numeric values, it is found that only ten_char_vector cannot be converted. I could not convert the character vector into numeric variables. 

## Problem 2

#### Loading up the penguin data set

```{r load penguin dataset}

library(palmerpenguins)
data("penguins", package = "palmerpenguins")

```

#### Finding the names and the important variables from the data set

The names in this data set are: 
`r names(penguins)`.

The unique *species* in the data set are: 
`r unique(penguins$species)`.

The unique *islands* in the data set are: 
`r unique(penguins$island)`.

The unique *years* in the data set are: 
`r unique(penguins$year, na.rm = TRUE)`.

The *sex* of the penguins in the data set are: 
`r unique(na.omit(penguins$sex))`.

The mean for *bill length* is `r round(mean(pull(penguins, bill_length_mm), na.rm = TRUE), digits = 2)`mm, while the standard deviation is `r round(sd(pull(penguins, bill_length_mm), na.rm = TRUE), digits = 2)`.

The mean for *bill depth* is `r round(mean(pull(penguins, bill_depth_mm), na.rm = TRUE), digits = 2)`mm, while the standard deviation is `r round(sd(pull(penguins, bill_depth_mm), na.rm = TRUE), digits=2)`.

The mean for *body mass* is `r round(mean(pull(penguins, body_mass_g), na.rm = TRUE), digits=2)`mm, while the standard deviation is `r round(sd(pull(penguins, body_mass_g), na.rm = TRUE), digits=2)`.

The mean for *flipper length* is `r round(mean(pull(penguins, flipper_length_mm),na.rm = TRUE))`mm, while the standard deviation is `r round(sd(pull(penguins, flipper_length_mm), na.rm = TRUE), digits=2)`.

There are `r nrow(penguins)` *rows* and `r ncol(penguins)` *columns* in this data set.

#### Creating a scatter plot

```{r create scatter plot, message = FALSE, warning = FALSE}

ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point() + ggtitle("Flipper versus Bill Length")

```

This is the scatter plot for Flipper versus Bill Length.

#### Saving the scatter plot

```{r save scatter plot, message = FALSE, warning = FALSE}

ggsave("penguin_scatter_plot.pdf", plot = last_plot(), height = 4, width = 6)

```

This saves the scatter plot as a pdf file names penguin_scatter_plot.