---
title: "p8105_hw1_ew2717"
author: "Eric Wang"
output: github_document
---

```{r include = FALSE}

library(tidyverse)
library(ggplot2)

```

## Problem 1



```{r creating data frame}

set.seed(10)

data_table <- tibble(
  my_variable = rnorm(10),
  logical_vec = my_variable > 0,
  ten_char_vector = c("c", "h", "a", "r", "a", "c", "t", "e", "r", "s"),
  three_levels = factor(c("one", "two", "three", "three", "two", "one", "one", "one", "three", "three"))
)

data_table
```

```{r finding mean of the variables}

mean(pull(data_table, my_variable))
mean(pull(data_table, logical_vec))
mean(pull(data_table, ten_char_vector))
mean(pull(data_table, three_levels))

```

The mean of my_variable and logical_vec can be calculated because they are numeric and logical vectors. ten_char_vector and three_levels, however, cannot have their means calculated because they are character and factor vectors.

```{r converting to variables to numeric, message = FALSE, warning = FALSE}

converted_num1 = as.numeric(pull(data_table, logical_vec))
converted_num2 = as.numeric(pull(data_table, ten_char_vector))
converted_num3 = as.numeric(pull(data_table, three_levels))

```

The logical and factor vectors can be converted to numeric values, while the character value cannot be converted. We are unable to convert characters or letters to numbers. For the logical vector, we are able to convert it to numeric because 1 can represent TRUE and 0 can represent FALSE. 

```{r finding converted mean}

mean(converted_num1)
mean(converted_num2)
mean(converted_num3)

```

After converting logical_vec, ten_char_vector, and three_levels to numeric values, it is found that only ten_char_vector cannot be converted. I could not convert the chracater vector into numeric variables. 

# Problem 2


```{r load penguin dataset}

library(palmerpenguins)
data("penguins", package = "palmerpenguins")

```

The names in this data set are: 
`r names(penguins)`.

The unique *species* in the data set are: 
`r unique(penguins$species)`.

The unique *islands* in the data set are: 
`r unique(penguins$island)`.

The unique *years* in the data set are: 
`r unique(penguins$year, na.rm = TRUE)`.

The *sex* of the penguins in the data set are: 
`r unique(na.omit(penguins$sex))`.

Important variables for *bill length* are shown in the table:

```{r, echo = FALSE}

summary(pull(penguins,bill_length_mm))
        
```

Important variables for *bill depth* are shown in the table:

```{r, echo = FALSE}

summary(pull(penguins,bill_depth_mm))
        
```


Important variables for *body mass* are shown in the table:

```{r, echo = FALSE}

summary(pull(penguins,body_mass_g))
        
```

Important variables for *flipper length* are shown in the table:

```{r, echo = FALSE}

summary(pull(penguins,flipper_length_mm))
        
```

As seen above, the mean flipper length is 200.9, which is rounded to `r round(mean(penguins$flipper_length_mm, na.rm = TRUE))`.

There are `r nrow(penguins)` *rows* and `r ncol(penguins)` columns in this data set.

```{r create scatter plot}

ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()
ggtitle("Flipper versus Bill Length")

```

This is the scatter plot for Flippter versus Bill Length.

```{r}

ggsave("penguin_scatter_plot.pdf", plot = last_plot(), height = 4, width = 6)

```

This saves the scatter plot as a pdf file names penguin_scatter_plot.